A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2018-09-29 paper/theor-experiment.tex
fixing spacing for one sided doc

\def \gunheight {25}
\def \quadheight {35}
\def \quadleft {127}
\def \kickerleft {160}
\def \leftarrow{10}
\def \bottomofarrows {29}
\def \topofarrows {34}
\begin{figure*}
	\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},x=1mm, y=1mm,]   
	\node (fig1) at (0,0)
	{\includegraphics[width=\textwidth]{Report/awa-drawing}};
	\node[fill=white, inner sep=2pt] (txt2) at (0,\gunheight) {\large Gun};
	\node[fill=white, inner sep=2pt] (txt2) at (20,\gunheight) {\large Accelerating Cavities};
	\node[fill=white, inner sep=2pt] (txt2) at (\quadleft,\quadheight) {\large Quadrupoles};
	\draw [blue, ->, line width=1] (\quadleft+\leftarrow,\topofarrows) -- (\quadleft+\leftarrow, \bottomofarrows);
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft,20) {\large $s_1$};
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft+8,20) {\large $s_2$};
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft,\quadheight) {\large Kicker};
	\draw [blue, ->, line width=1] (\kickerleft+\leftarrow-3,\topofarrows) -- (\kickerleft+\leftarrow-3, \bottomofarrows);	
	\draw [black, ->, line width=1] (0,-2) -- (100, -2);
	\node[fill=white, inner sep=2pt] (txt2) at (50,0) {\large Beam Direction};
	\end{tikzpicture}
	\caption{Side view of the high charge linac at the AWA. 
		All hardware in this drawing is currently installed. 
	Note locations $s_1$ and $s_2$, before and after the kicker.}
	\label{awa-linac}
	\vspace{-7em}
\end{figure*}

\def \septheight {65}
\def \bottomarrow {57}
\def \kleft {13}
\def \sleft {41.5}
\def \sheight {47}
\def \beam {35}
\begin{figure*}
	\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},x=1mm, y=1mm,]   
	\node (fig2) at (0,0)
	{\includegraphics[width=\textwidth]{Report/tba_dogleg}};
	\node[fill=white, inner sep=2pt] (txt2) at (3,\sheight) {\large $s_1$};
	\node[fill=white, inner sep=2pt] (txt2) at (20,\sheight) {\large $s_2$};
	\node[fill=white, inner sep=2pt] (txt2) at (46,45) {\large $s_3$};
	\node[fill=white, inner sep=2pt] (txt2) at (33,\septheight) {\large Septum};
	\draw [blue, ->, line width=1] (\sleft,\septheight) -- (\sleft, \bottomarrow);	
	\node[fill=white, inner sep=2pt] (txt2) at (7,\septheight) {\large Kicker};
	\draw [blue, ->, line width=1] (\kleft,\septheight) -- (\kleft, \bottomarrow);
	\draw [black, ->, line width=1] (0,\beam) -- (100,\beam);
	\node[fill=white, inner sep=2pt] (txt2) at (50,\beam+2) {\large Beam Direction};
	\end{tikzpicture}
	\vspace{-12em}
	\caption{Continuation of the high charge beam line layout at the AWA, top view. 
		This is the proposed two beam acceleration section. 
		Only the kicker in this drawing is installed. 
	Note $s_3$, the entrance to the fifth quadrupole on the beam line.
	This is the optimization location.}
	\label{awa-tba}
\end{figure*}
%
%\vspace{-1em}

@@

In the drifts and linac tanks, $dT=\num{1E-11}$ was sufficient. 
However, it was not acceptable near the quadrupoles. 
For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but we see discrepancies in the transverse 
(rms$_x$ and $\epsilon_x$) for low fidelity results. This discrepancy is 
what led to the decision to adjust the time steps w.r.t. beam line elements. 
In the linac and drift sections $dT=\num{1E-11}$ was used. 
Near sensitive elements such as the quadrupoles, kicker, and septum, 
a time step of $dT=\num{1E-12}$ was used.
The resulting simulations are low fidelity in most places, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig.~\ref{tstep}.  Mid fidelity simulations used steps of  $dT=\num{1E-12}$ everywhere.
The average run time of each simulation with the adjusted time steps was 1.6~minutes.
In comparison, the mid fidelity simulation ran for 18~minutes.
Note, a smaller time step, \num{1E-13}, is always used in the gun where the 
beam has low energy and changing rapidly.
%A summary of the parameters used is listed in Table \ref{fidelity}.

\begin{figure}%[h]
	\centering
	\includegraphics[width=\linewidth]{Report/timestep_comparison_paper}
	\caption{Comparison of different fidelity models (dT stands for time step).}
	\label{tstep}
\end{figure}

\iffalse
\begin{table}%[h!]
	\begin{center}
		\caption{Low and mid fidelity simulation parameters.}
		\label{fidelity}
		\begin{tabular*}{0.48\textwidth}{l @{\extracolsep{\fill}} C c D }
			\hline\noalign{\smallskip}
			& Number of particles & dT (seconds) & Time to simulation (minutes)\\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			low fidelity  			&  10,000   & $5 \times10^{-11}$  &  2.6 \\
			adjusted low fidelity  	&  10,000   & $2 \times 10^{-12}$, $1 \times 10^{-11}$ & 1.6 \\
			mid fidelity 			&  100,000  & $1 \times 10^{-12}$ &  18\\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}
\fi

@@

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty four hours and
Pareto fronts at $s_1$ and $s_2$, see Fig.~\ref{awa-linac}.
From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated only 53 generations 
compared to the experiment with the maximum number, ex-1 at 96 generations.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto}, we can see this is not the case.
Similar arguments can be made for ex-2, which evaluated about 15 less generations.
The Pareto fronts at $s_2$, are nearly identical. We could expect
this trend would continue given more time. 
When looking at the Pareto front at $s_1$, only ex-4 has a slightly 
larger range compared to the others.
With ex-2 and ex-3 eliminated due to evaluation time, 
and a slightly better Pareto front at $s_1$ for ex-4, 
the hyper parameters in ex-4 were chosen as the default values for subsequent runs.

\begin{figure}
	\centering
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto1}
		(a) Pareto fronts for ex-1 through ex-4 at $s_1$.
		%\caption{(a) Pareto fronts at $s_1$.}
		%\label{expareto1}
		\vspace{2em}
	\end{minipage}
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto2}
		(b) Pareto fronts for ex-1 through ex-4 at $s_2$.
		%\caption{(b) Pareto fronts at $s_2$.}
		%\label{expareto2}
	\end{minipage}
	\caption{Comparison of Pareto fronts for initial optimization experiments, ex-1 through ex-4.}
	\label{expareto}
\end{figure}

>>>>

\def \gunheight {25}  %25} %double sided number
\def \quadheight {35}
\def \quadleft {115}  %127} %double sided number
\def \kickerleft {145}  %160} % two sided number
\def \leftarrow{10}
\def \bottomofarrows {29}
\def \topofarrows {34}
\begin{figure*}
	\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},x=1mm, y=1mm,]   
	\node (fig1) at (0,0)
	{\includegraphics[width=\textwidth]{figures/awa-drawing}};
	\node[fill=white, inner sep=2pt] (txt2) at (0,\gunheight) {\large Gun};
	\node[fill=white, inner sep=2pt] (txt2) at (20,\gunheight) {\large Accelerating Cavities};
	\node[fill=white, inner sep=2pt] (txt2) at (\quadleft-5,\quadheight) {\large Quadrupoles};
	\draw [blue, ->, line width=1] (\quadleft+\leftarrow,\topofarrows) -- (\quadleft+\leftarrow, \bottomofarrows);
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft,20) {\large $s_1$};
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft+8,20) {\large $s_2$};
	\node[fill=white, inner sep=2pt] (txt2) at (\kickerleft,\quadheight) {\large Kicker};
	\draw [blue, ->, line width=1] (\kickerleft+\leftarrow-3,\topofarrows) -- (\kickerleft+\leftarrow-3, \bottomofarrows);	
	\draw [black, ->, line width=1] (0,-2) -- (100, -2);
	\node[fill=white, inner sep=2pt] (txt2) at (50,0) {\large Beam Direction};
	\end{tikzpicture}
	\caption{Side view of the high charge linac at the AWA. 
		All hardware in this drawing is currently installed. 
	Note locations $s_1$ and $s_2$, before and after the kicker.}
	\label{awa-linac}
	%\vspace{-7em}
\end{figure*}

\def \septheight {65}
\def \bottomarrow {55} %57}
\def \kleft {13}
\def \sleft {38}%41.5}
\def \sheight {45}%47}
\def \beam {35}
\begin{figure*}
	\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},x=1mm, y=1mm,]   
	\node (fig2) at (0,0)
	{\includegraphics[width=\textwidth]{figures/tba_dogleg}};
	\node[fill=white, inner sep=2pt] (txt2) at (3,\sheight) {\large $s_1$};
	\node[fill=white, inner sep=2pt] (txt2) at (20,\sheight) {\large $s_2$};
	\node[fill=white, inner sep=2pt] (txt2) at (41,40) {\large $s_3$}; %(46,45)
	\node[fill=white, inner sep=2pt] (txt2) at (33,\septheight) {\large Septum};
	\draw [blue, ->, line width=1] (\sleft,\septheight) -- (\sleft, \bottomarrow);	
	\node[fill=white, inner sep=2pt] (txt2) at (7,\septheight) {\large Kicker};
	\draw [blue, ->, line width=1] (\kleft,\septheight) -- (\kleft, \bottomarrow);
	\draw [black, ->, line width=1] (0,\beam) -- (100,\beam);
	\node[fill=white, inner sep=2pt] (txt2) at (50,\beam+2) {\large Beam Direction};
	\end{tikzpicture}
	\vspace{-12em}
	\caption{Continuation of the high charge beam line layout at the AWA, top view. 
		This is the proposed two beam acceleration section. 
		Only the kicker in this drawing is installed. 
	Note $s_3$, the entrance to the fifth quadrupole on the beam line.
	This is the optimization location.}
	\label{awa-tba}
\end{figure*}
%
%\vspace{-1em}

@@

In the drifts and linac tanks, $dT=\num{1E-11}$ was sufficient. 
However, it was not acceptable near the quadrupoles. 
For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but discrepancies are seen in the transverse 
(rms$_x$ and $\epsilon_x$) for low fidelity results. This discrepancy is 
what led to the decision to adjust the time steps w.r.t. beam line elements. 
In the linac and drift sections $dT=\num{1E-11}$ was used. 
Near sensitive elements such as the quadrupoles, kicker, and septum, 
a time step of $dT=\num{1E-12}$ was used.
The resulting simulations are low fidelity in most places, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig.~\ref{tstep}.  Mid fidelity simulations used steps of  $dT=\num{1E-12}$ everywhere.
The average run time of each simulation with the adjusted time steps was 1.6~minutes.
In comparison, the mid fidelity simulation ran for 18~minutes.
Note, a smaller time step, \num{1E-13}, is always used in the gun where the 
beam has low energy and changing rapidly.
%A summary of the parameters used is listed in Table \ref{fidelity}.

\begin{figure}%[h]
	\centering
	\includegraphics[width=\linewidth]{figures/timestep_comparison_paper}
	\caption{Comparison of different fidelity models (dT stands for time step).}
	\label{tstep}
\end{figure}

@@

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty four hours and
Pareto fronts at $s_1$ and $s_2$, see Fig.~\ref{awa-linac}.
From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated only 53 generations 
compared to the experiment with the maximum number, ex-1 at 96 generations.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto}, but this is not the case.
Similar arguments can be made for ex-2, which evaluated about 15 less generations.
The Pareto fronts at $s_2$, are nearly identical. It is expected
this trend would continue given more time. 
When looking at the Pareto front at $s_1$, only ex-4 has a slightly 
larger range compared to the others.
With ex-2 and ex-3 eliminated due to evaluation time, 
and a slightly better Pareto front at $s_1$ for ex-4, 
the hyper parameters in ex-4 were chosen as the default values for subsequent runs.

\begin{figure}
	\centering
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figures/ex-pareto1}
		(a) Pareto fronts for ex-1 through ex-4 at $s_1$.
		%\caption{(a) Pareto fronts at $s_1$.}
		%\label{expareto1}
		\vspace{2em}
	\end{minipage}
	\begin{minipage}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{figures/ex-pareto2}
		(b) Pareto fronts for ex-1 through ex-4 at $s_2$.
		%\caption{(b) Pareto fronts at $s_2$.}
		%\label{expareto2}
	\end{minipage}
	\caption{Comparison of Pareto fronts for initial optimization experiments, ex-1 through ex-4.}
	\label{expareto}
\end{figure}
